```{r lockfile, include=FALSE}
renv::use(
  base64enc    = "base64enc@0.1-6",
  bit          = "bit@4.6.0",
  bit64        = "bit64@4.8.2",
  bslib        = "bslib@0.11.0",
  cachem       = "cachem@1.1.0",
  cli          = "cli@3.6.6",
  clipr        = "clipr@0.8.1",
  cpp11        = "cpp11@0.5.5",
  crayon       = "crayon@1.5.3",
  digest       = "digest@0.6.39",
  dplyr        = "dplyr@1.2.1",
  evaluate     = "evaluate@1.0.5",
  farver       = "farver@2.1.2",
  fastmap      = "fastmap@1.2.0",
  fontawesome  = "fontawesome@0.5.3",
  fs           = "fs@2.1.0",
  generics     = "generics@0.1.4",
  ggplot2      = "ggplot2@4.0.3",
  glue         = "glue@1.8.1",
  gtable       = "gtable@0.3.6",
  highr        = "highr@0.12",
  hms          = "hms@1.1.4",
  htmltools    = "htmltools@0.5.9",
  isoband      = "isoband@0.3.0",
  jquerylib    = "jquerylib@0.1.4",
  jsonlite     = "jsonlite@2.0.0",
  knitr        = "knitr@1.51",
  labeling     = "labeling@0.4.3",
  lifecycle    = "lifecycle@1.0.5",
  magrittr     = "magrittr@2.0.5",
  memoise      = "memoise@2.0.1",
  mime         = "mime@0.13",
  pillar       = "pillar@1.11.1",
  pkgconfig    = "pkgconfig@2.0.3",
  prettyunits  = "prettyunits@1.2.0",
  progress     = "progress@1.2.3",
  purrr        = "purrr@1.2.2",
  R6           = "R6@2.6.1",
  rappdirs     = "rappdirs@0.3.4",
  RColorBrewer = "RColorBrewer@1.1-3",
  readr        = "readr@2.2.0",
  renv         = "renv@1.2.3",
  rlang        = "rlang@1.2.0",
  rmarkdown    = "rmarkdown@2.31",
  S7           = "S7@0.2.2",
  sass         = "sass@0.4.10",
  scales       = "scales@1.4.0",
  stringi      = "stringi@1.8.7",
  stringr      = "stringr@1.6.0",
  tibble       = "tibble@3.3.1",
  tidyr        = "tidyr@1.3.2",
  tidyselect   = "tidyselect@1.2.1",
  tinytex      = "tinytex@0.60",
  tzdb         = "tzdb@0.5.0",
  utf8         = "utf8@1.2.6",
  vctrs        = "vctrs@0.7.3",
  viridisLite  = "viridisLite@0.4.3",
  vroom        = "vroom@1.7.1",
  withr        = "withr@3.0.3",
  xfun         = "xfun@0.59",
  yaml         = "yaml@2.3.12"
)
```

# Primary DWI Biomarker {#sec-dwi-biomarker1}

```{r setup}
library(dplyr)
library(ggplot2)
library(tidyr)
library(readr)
library(fs)
```

In @sec-bedpostx, we covered the outputs of BedpostX, which is dmethod for modeling how water diffuses in each voxel and estimating their implied fiber orientation distributions. The outputs of BedpostX were used as the inputs for one of the primary A2CPS biomarker, which is based on a finding reported by @vachon2016corticolimbic: as compared with patients whose subacute back pain recovered, those whose pain persisted exhibited higher connectivity at baseline within a corticolimbic subnetwork involving the amygdala, nucleus accumbens, and the dorsal medial prefrontal cortex (Am-NAc-mPFC). This kit describes the derivatives produced while extracting that biomarker.

## Starting Project

### Locate Data

{{< include _snippets/mri-location.qmd >}}

The derivatives for this biomarker are a collection of three kinds of outputs:

```{bash}
#| eval: false
$ ls derivatives/dwi_biomarker1/
move_masks  networks  probtrackx
```

Each of these are organizing in a "BIDS-ish" manner.

```{bash}
#| eval: false
$ tree derivatives/dwi_biomarker1/move_masks/sub-10003
derivatives/dwi_biomarker1/move_masks/sub-10003
└── ses-V1
    └── dwi
        ├── sub-10003_ses-V1_masks
        ├── sub-10003_ses-V1_space-dwi_desc-modules_dseg.nii.gz
        ├── sub-10003_ses-V1_space-dwifslstd_desc-module0_dseg.nii.gz
        ├── sub-10003_ses-V1_space-dwifslstd_desc-module1_dseg.nii.gz
        └── sub-10003_ses-V1_space-dwifslstd_desc-module2_dseg.nii.gz
2 directories, 5 files

$ tree derivatives/dwi_biomarker1/networks/sub-10003
derivatives/dwi_biomarker1/networks/sub-10003
└── ses-V1
    └── dwi
        └── network_summaries.tsv
2 directories, 1 file

$ tree derivatives/dwi_biomarker1/probtrackx/sub-10003
derivatives/dwi_biomarker1/probtrackx/sub-10003
└── ses-V1
    └── dwi
        ├── coords_for_fdt_matrix1
        ├── fdt_matrix1.dot.gz
        ├── fdt_paths.nii.gz
        ├── probtrackx.log
        └── waytotal
2 directories, 5 files
```

- move_masks
  - The biomarker is calculated in the native dwi space. These files contain information about
    - `*space-dwi_desc-modules_dseg.nii.gz`: the relevant masks in participant-native dwi space. All three subnetworks are contained as non-overlapping clusters.
    - `*space-dwifslstd_desc-module*_dseg.nii.gz`: Binary mask files for each of the three subnetworks.
    - `*masks` a text file pointing to each of the binary mask files that is used as input for the `probtrackx` step.
- probtrackx
  - Results from probabilistic tract tracing (with [ProbtrackX](https://fsl.fmrib.ox.ac.uk/fsl/docs/diffusion/probtrackx.html)).
- networks
  - Tabular information summarizing connectivity within the networks.


Most users will be interested in the tabular information. A data dictionary is available at the top-level of the `network` folder (`derivatives/dwi_biomarker1/network_summaries.json`). Note that we use the terms "subnetwork" and "module" interchangeably. 

```{json}
{
    "Deg_mean": {
        "LongName": "Mean Degree",
        "Description": "Average nodal degree within module (after thresholding network to 10% density)",
        "TermURL": "https://networkx.org/documentation/stable/reference/classes/generated/networkx.Graph.degree.html"
    },
    "Density": {
        "LongName": "Density",
        "Description": "Density of the module (after thresholding network to 10% density)",
        "TermURL": "https://networkx.org/documentation/stable/reference/generated/networkx.classes.function.density.html"
    },
    "WM_connections": {
        "LongName": "White Matter Connections",
        "Description": "A summary of the total connections in the network: sum(Degs) / (nvox * (nvox - 1) / 2)"
    },
    "ModuleNumber": {
        "LongName": "Module Number",
        "Description": "Index of the network module"
    },
    "ModuleName": {
        "LongName": "Module Name",
        "Description": "Name of the module from the reference (https://doi.org/10.1093/brain/aww100)",
        "Levels": {
            "mPFCventral-amyg": "ventral medial PFC–amygdala",
            "OFC-amyg-hipp": "orbitofrontal cortex–amygdala–hippocampus",
            "mPFCdorsal-amyg-NAc": "dorsal medial PFC–amygdala–nucleus accumbens",
            "WholeNetwork": "Full network (not a module)"
        }
    },
    "IsBiomarker": {
        "LongName": "Biomarker Indicator",
        "Description": "Boolean indicating whether this module is a primary biomarker for A2CSP"
    },
    "sub": {
        "LongName": "Subject",
        "Description": "Study Participant, BIDS Subject ID",
        "TermURL": "https://bids-specification.readthedocs.io/en/v1.9.0/appendices/entities.html#sub"
    },
    "ses": {
        "LongName": "Session",
        "Description": "Visit, Protocol, BIDS Session ID",
        "Levels": {
            "V1": "baseline_visit",
            "V3": "3mo_postop"
        },
        "TermURL": "https://bids-specification.readthedocs.io/en/v1.9.0/appendices/entities.html#ses"
    }
}
```

### Extract Data

The tables always have the same schema and so can be loaded together using standard tools.

```{r networks}

networks <- read_tsv(fs::dir_ls(
  "data/dwi_biomarker1/networks",
  glob = "*tsv",
  recurse = TRUE
))

head(networks)
```

For understanding these, it may help to plot all available metrics.

::: {#fig-dwi-p-metrics}

```{r dwi-p-metrics}
networks |>
  pivot_longer(c(Deg_mean, Density, WM_connections), names_to = "metric") |>
  ggplot(aes(y = ModuleName, x = value)) +
  facet_wrap(~metric, scales = "free_x") +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(alpha = 0.2)
```

Distribution of Structural Connectivity Metrics Related to Primary dMRI Biomarker. Deg_mean: Average degree within network. Density: Total degree of network divided by the number of nodes in the network. WM_connections: 2x the density. Note that these are calculated after thresholding the WholeNetwork density to 10%.

:::

In @fig-dwi-p-metrics, we can see that the thresholding did not affect most connectivity matrices: the density of the whole network was already below 10%. Nevertheless, there is variability in the subnetworks that may have predictive value. 

## Considerations While Working on the Project

### Data Generation

The pipeline for this connectivity biomarker was extracted by first passing outputs of BEDPOSTX to FSL’s probabilistic tractography routine PROBTRACKX2. When conducting the tracking, the following options were used: a loop check, a one-way condition, a curvature threshold of 0.2, 2000 steps, a step length of 0.5, 5000 samples, a 0.01 volume fraction threshold for consideration of subsidiary fiber orientations, a distance threshold of 0, no random sampling around the voxel, modified Euler streamlining, and correction of path distributions for path length. Note that the probtrackx command is present in a log file for each participant (e.g., `derivatives/dwi_biomarker1/probtrackx/sub-10003/probtrackx.log`).

As a seed for probabilistic tractography, we used a voxel-wise corticolimbic mask from @vachon2016corticolimbic. This mask comprised three subnetworks: the Am-NAc-mPFC, a ventral mPFC-Amygdala network, and one involving the amygdala, hippocampus, and orbitofrontal cortex. The resulting connectivity matrix was symmetrized and binarized (thresholded to a density of 10%). The final biomarker was the total degree of the nodes within the subnetwork divided by the maximum possible number of connections in the subnetwork. This connectivity analysis was performed in each subject’s native space (1.7 mm^3^); the corticolimbic mask was transformed into native space using the warps estimated by QSIPrep during preprocessing, with ANTs (nearest-neighbor interpolation). 

### Interpretation of "Streamlines"

The information provided by probabilistic tracktography is often misrepresented. If you are brand new to the subject, the [official FSL tutorial on PROBTRACKX](https://www.youtube.com/watch?v=_N8hzzo818U) is recommended. In particular, keep in mind is that, although we often want to know information about the "strength" or "integrity" of connections, ProbtrackX only provides information about whether a connection between two regions is likely; more streamlines correspond to higher chances of there being a connection, and not anything about the quality or size of those connections. There is almost certainly a relationship between whether ProbtrackX determines a connection to be likely and the strength of a connection, with strength meaning something like the capacity for information flow, or, more concretely, the number of dendrites comprising a track. But the nature of that relationship is unknown, and caution should be exercised when describing these outputs.

### Calculating Additional Metrics

In some analyses, it may be interesting to go back to the connectivity graphs as output by ProbtrackX (e.g., to explore alternative thresholds). In that case, you would need to start with the files in `derivatives/dwi_biomarker1/probtrackx`. Example code for working with those files is available [here](https://github.com/a2cps/biomarker-extractor/blob/8c6868d95ba94fd5ef6c82c4cc1266f42ea2e9d3/src/biomarkers/flows/dwi_biomarker1.py). The key point is that ProbtrackX stores the connectivity outputs in an asymmetric [Coordinate List format](https://en.wikipedia.org/wiki/Sparse_matrix#Coordinate_list_(COO)) (the asymmetry implies that there are entries for both source -> target and target -> source, even though the connections themselves are undirected). 

### Variability Across Scanners

{{< include _snippets/mri-scanner-variability.qmd >}}

### Data Quality

{{< include _snippets/mri-qc.qmd >}}

### Data Generation

These outputs were generated by the [dwi_biomarker1_app](https://github.com/a2cps/mri_imaging_pipeline/tree/master/dwi_biomarker1_app). 

### Citations

If you use these products in your analyses, please cite the relevant papers from [FSL](https://fsl.fmrib.ox.ac.uk/fsl/docs/diffusion/probtrackx.html). Additionally, please cite the source of this biomarker [@vachon2016corticolimbic].

{{< include _snippets/a2cps_citations.qmd >}}

When using neuroimaging derivatives, please also cite @sadil_acute_2024.
