# Manual Derivative QC {#mri-deriv-qc}

```{r setup}
library(dplyr)
library(duckplyr)
library(purrr)
library(ggplot2)
```

As of Release 3.0, a subset of the neuroimaging derivatives have undergone manual QC. 
This kit shows how the ratings are packaged.

## Starting Project


{{< include _snippets/mri-location.qmd >}}

The Derivative QC metrics are underneath `derivatives/qc`. They consist of four tables

```{bash}
#| eval: false
$ ls mris/derivatives/qc/
coordinate.parquet  image.parquet  rating.parquet  session.parquet
```

### Extract Data

First, let's look through the schema of each table.

#### session

This table contains information about the rating session (e.g., helpful for determining rater consistency).

```{r session}
read_parquet_duckdb("data/qc/session.parquet") |>
  head()
```

- `id`: ID of session
- `step`: ID indicating the kind of target being rated (https://github.com/psadil/django-qcapp-ratings/blob/1080ae40304c067fec283fc04bd956834c61d99c/src/django_qcapp_ratings/models.py#L9-L14)
    - MASK = 0
    - SPATIAL_NORMALIZATION = 1
    - SURFACE_LOCALIZATION = 2
    - FMAP_COREGISTRATION = 3
    - DTIFIT = 4
- `user`: ID user who was doing to rating (note: user id capturing currently exhibits some errors)
- `created`: datetime stamp of when the session started

#### image

This table contains information about the to-be rated targets. 

```{r image}
read_parquet_duckdb("data/qc/image.parquet") |>
  select(-img) |> # just for display purposes, so this webpage is not too big
  head()
```

- `id`: ID of the to-be-rated image
- `img`: binary large object containing the image
- `slice`: indicator for which part of the image was displayed
- `file1`: the main source file from which the image was derived (e.g., the name of the mask)
- `file2`: supplementary file used to make the to-be-rated image (e.g., anatomical underlay)
- `display`: ID indicating the orientation of the image (https://github.com/psadil/django-qcapp-ratings/blob/1080ae40304c067fec283fc04bd956834c61d99c/src/django_qcapp_ratings/models.py#L23-L26)
    - X = 0
    - Y = 1
    - Z = 2
- `step`: ID indicating the kind of target (https://github.com/psadil/django-qcapp-ratings/blob/1080ae40304c067fec283fc04bd956834c61d99c/src/django_qcapp_ratings/models.py#L9-L14)
    - MASK = 0
    - SPATIAL_NORMALIZATION = 1
    - SURFACE_LOCALIZATION = 2
    - FMAP_COREGISTRATION = 3
    - DTIFIT = 4
- `created`: when the image was created (helpful for determining version of the release)

#### rating

This table contains the ratings given as "overal" scores. They apply to ratings for targets fmap coregistration and dtifit.

```{r rating}
read_parquet_duckdb("data/qc/rating.parquet") |>
  head()
```

- `id`: ID of each rating
- `source_data_issue`: boolean indicating whether the reviewer noticed an issue with the raw data quality
- `rating`: the rating provided (https://github.com/psadil/django-qcapp-ratings/blob/1080ae40304c067fec283fc04bd956834c61d99c/src/django_qcapp_ratings/models.py#L17-L20)
    - PASS = 0
    - UNSURE = 1
    - FAIL = 2
- img_id: ID of the image that was rated (see image.parquet)
- session: ID of the rating session (see session.parquet)
- notes: freeform text used to provide details about the rating
- created: datetime stamp of when the rating was produced

#### coordinate

This table contains the coordinates that were clicked during reviews, locations indicating some source data issue. These were used to rate the mask, spatial normalization, and surface localization.

```{r coordinate}
read_parquet_duckdb("data/qc/coordinate.parquet") |>
  head()
```

- `id`: ID of each clicked coordinate
- `source_data_issue`: boolean indicating whether the reviewer noticed an issue with the raw data quality
- `x`: horizontal coordinate of the target issue
- `y`: vertical coordinate of the target issue
- `img_id`: ID of the image that was rated (see image.parquet)
- `session`: ID of the rating session (see session.parquet)
- `notes`: freeform text used to provide details about the rating
- `created`: datetime stamp of when the rating was produced

Note: rows where the x and y coordinates are NA indicate that the reviewer did not observe any issues with the target.


### Example Image Plotting

Let's see how we can display clicks on an image.

```{r}

# grab images associated with spatial normalization
spatial_norm_images <- read_parquet_duckdb("data/qc/image.parquet") |>
  filter(step == 1, display == 0) |>
  select(id, img)

# grab image with lots of clicks
img_with_many_clicks <- read_parquet_duckdb("data/qc/coordinate.parquet") |>
  filter(!is.na(x)) |>
  semi_join(spatial_norm_images, by = join_by(img_id == id)) |>
  summarize(n_clicks = n(), .by = c(img_id)) |>
  slice_head(n = 1, by = n_clicks)
img_with_many_clicks
```

We're going to need a helper function to convert the img BLOBs into a tibble for display with ggplot.

```{r helper}
to_tbl <- function(.i) {
  apply(.i, 2, rgb) |>
    as_tibble() |>
    mutate(y = row_number()) |>
    tidyr::pivot_longer(-y, names_to = "x") |>
    mutate(
      x = stringr::str_extract(x, "[[:digit:]]+") |> as.integer(),
      y = max(y) - y + 1,
      rgb = map(value, col2rgb),
      r = map_dbl(rgb, pluck, 1),
      g = map_dbl(rgb, pluck, 2),
      b = map_dbl(rgb, pluck, 3)
    )
}
```

Now, return to the image.parquet file, do a filter join to get just a single row, select it, apply the helper, and plot.

```{r showimg}
# now show the associated blob
read_parquet_duckdb("data/qc/image.parquet") |>
  semi_join(img_with_many_clicks, by = join_by(id == img_id)) |>
  pluck("img", 1) |>
  png::readPNG() |>
  to_tbl() |>
  ggplot(aes(x = x, y = y)) +
  terrainr::geom_spatial_rgb(aes(r = r, g = g, b = b)) +
  coord_fixed() +
  theme_void()
```

By referring back to img.parquet file, we can see which file was plotted.

```{r imgmeta}
read_parquet_duckdb("data/qc/image.parquet") |>
  semi_join(img_with_many_clicks, by = join_by(id == img_id)) |>
  select(-img)
```

### Use of Ratings to Exclude Images


```{r}
# grab all fieldmap images
fmap_coregistration_images <- read_parquet_duckdb("data/qc/image.parquet") |>
  filter(step == 3) |>
  select(id, file1, file2, display)

# search for all coregistrations that were marked as having failed
definite_coregistration_failures <- read_parquet_duckdb(
  "data/qc/rating.parquet"
) |>
  inner_join(fmap_coregistration_images, by = join_by(img_id == id)) |>
  filter(rating == 2) |>
  select(file1, file2, display) |>
  # pluck out useful metadata from the image
  mutate(
    sub = stringr::str_extract(file1, "(?<=sub-)[[:digit:]]{5}") |>
      as.integer(),
    task = stringr::str_extract(file1, "cuff|rest"),
    run = stringr::str_extract(file1, "(?<=run-0)[1|2]")
  ) |>
  distinct(sub, task, run) |>
  collect()

```

```{r}
nrow(definite_coregistration_failures)
```

```{r}
definite_coregistration_failures |>
  head()
```

## Considerations While Working on the Project

### Data Generation

Quality control was done with the A2CPS Derivative Image Review Tool (DIRT). This application is actively under development here: https://github.com/psadil/django-qcapp-ratings/tree/main/src/django_qcapp_ratings.

### Citations

If you use these results, please cite the MRIQC report [@esteban_mriqc_2017] and all relevant citations of the pipeline configuration. 

{{< include _snippets/a2cps_citations.qmd >}}

When using neuroimaging derivatives, please also cite @sadil_acute_2024.
