# NDA Download Example, MRI {#sec-nda-mri}

This page describes the A2CPS rawdata as downloaded from the NDA. 

If you do not yet have access to the NDA and plan to download data from the NDA, please follow the instructions on the A2CPS For Researchers page ([Accessing our Data](https://a2cps.org/researchers/accessing-our-data/)) and the linked tutorials.

## Data Overview

This starter kit will assume that the data have been downloaded using [nda-tools](https://github.com/NDAR/nda-tools). In this example, a Data Package has been created that includes the rawdata. You will see a list of files like the following

```bash
$ tree -L 2
.
├── dataset_collection.txt
├── datastructure_manifest.txt
├── experiments
│   ├── experiment_2652
│   └── experiment_2655
├── image03
│   ├── corral-secure
│   └── manifest-files
├── image03.txt
├── md5_values.txt
├── package_file_metadata_1235008.txt.gz
├── package_info.txt
└── README.pdf
```

## Data Structures

This is raw imaging data, and so it is stored in the [image03 Data Structure](https://nda.nih.gov/data-structure/image03) (link goes to data dictionary). Metadata is stored in the `image03.txt`. This is where you will find information such as participant age, participant sex, and scan acquisition date. 

Note that, as with downloads of all NDA data structures, the second line of this table includes a partial data dictionary. That is, the header is on the first line, and the first row of data is on the third line.

## Imaging Files

The imaging files are underneath `image03`. 

```bash
$ ls image03/corral-secure/projects/A2CPS/products/consortium-data/pre-surgery/mris/bids/ | head
dataset_description.json
README
sessions.json
sub-10003
sub-10005
sub-10008
sub-10010
sub-10011
sub-10013
```

The contents of that folder match the contents described in @sec-raw-mri (note that the aggregated table `participants.tsv` is not included in the NDA download, as the information is available in `image03.txt`).

## Experiments

The rawdata folder includes an “experiments“ folder, which provides some information on how the two types of fMRI scans were collected.

```{r}
#| echo: false
tibble::tribble(
  ~"Experiment", ~"Files",
  "experiment_2652 (rfMRI)", '{{< downloadthis files/experiment_2652.xml dname="experiment_2652" label="experiment_2652.xml" >}} {{< downloadthis files/A2CPS_rest.txt dname="A2CPS_rest" label="A2CPS_rest.txt" >}}',
  "experiment_2655 (tfMRI)", '{{< downloadthis files/experiment_2655.xml dname="experiment_2655" label="experiment_2655.xml" >}} {{< downloadthis files/A2CPS_cuff.txt dname="A2CPS_cuff" label="A2CPS_cuff.txt" >}}'
)
```


## Additional NDA Metadata

The rest of the files are metadata associated with the storage of the dataset in the NDA or information about the Data Package. Most users will not need to interact with these files unless something has gone wrong with the download.

```{r}
#| echo: false
tibble::tribble(
  ~`File/Directory`, ~`Description/Notes`,
  '{{< downloadthis files/dataset_collection.txt dname="dataset_collection" label="dataset_collection.txt" >}}', "Table with identifiers relating to how the data were uploaded",
  '{{< downloadthis files/datastructure_manifest.txt dname="datastructure_manifest" label="datastructure_manifest.txt" >}}', "Table with partial information about how the files are stored on the NDA AWS S3 bucket. This example file contains the first 10 rows.",
  '{{< downloadthis files/md5_values.txt dname="md5_values" label="md5_values.txt" >}}', "Table with MD5 File Checksums for each of the files in the image03 folder. This example file contains the first 10 rows.",
  '{{< downloadthis files/package_file_metadata_1235008.txt dname="package_file_metadata_1235008" label="package_file_metadata_1235008.txt" >}}', "Additional metadata related to the location of imaging files within the NDA and their S3 bucket",
  '{{< downloadthis files/package_info.txt dname="package_info" label="package_info.txt" >}}', "High-level information about the Data Package and Filter Cart at the time of Data Package creation. Note that the counts listed in this file may not relate to the files that were downloaded (e.g., this file refers to 1053 participants in the filter, but 585 participants were downloaded).",
  '{{< downloadthis files/README.pdf dname="README" label="README.pdf" >}}', "Information that pertains to some NDA Data Packages."
)
```

## Contacts

- slack: [#help-imaging](https://a2cps-pain.slack.com/archives/C07SBJJLFMG)
- email:
  - [James Ford](https://a2cps.atlassian.net/wiki/people/712020:b74bac07-188d-4937-94b0-e97a7d5ea96e?ref=confluence)
  - [Patrick Sadil](https://a2cps.atlassian.net/wiki/people/712020:9b1f0b7e-3a1b-4955-856b-f0b1a07e458e?ref=confluence)

