# ICA Connectivity {#sec-postgift}

```{r setup}
library(readr)
library(dplyr)
library(ggplot2)
library(duckplyr)
library(fs)
library(tidyr)
library(ggdist)
```

The @sec-gift products provide outputs in a raw format that will be familiar to users of that library. To facilitate analyses, A2CPS has repackaged those products into a tabular format. This kit covers the repackaged structure.

## Starting Project

### Locate Data

{{< include _snippets/mri-location.qmd >}}

The MRIQC metrics are underneath `derivatives/postgift`. Let's take a look.

```{bash}
#| eval: false
$ ls postgift
amplitude  amplitude.json  biomarkers  biomarkers.json  connectivity  connectivity.json
```

### Extract Data

The tabular data comprise [parquet files](https://parquet.apache.org/) that have been partitioned in a [hive style](https://hive.apache.org/). Each table is a different summary of the components derived from GIFT.

- [amplitude](https://github.com/a2cps/snapshot/blob/main/src/snapshot/data/gift_amplitude.json): a measure of component amplitudes. In particular, they are the fractional Amplitude of Low Frequency Fluctuations [@zou_improved_2008].
- [connectivity](https://github.com/a2cps/snapshot/blob/main/src/snapshot/data/gift_connectivity.json): the (product-moment) correlations between component timecourses.
- [biomarkers](https://github.com/a2cps/snapshot/blob/main/src/snapshot/data/gift_biomarkers.json): putative biomarkers of chronic pain calculated for A2CPS analyses.

Let's take a look at just the amplitudes.

```{r loadfalff}

amplitude <- read_parquet_duckdb(
  dir_ls(
    "data/postgift/amplitude",
    recurse = TRUE,
    glob = "*parquet"
  ),
  prudence = "lavish"
)

head(amplitude)
```

To get a feel for the data, let's see whether there are differences between the amplitudes in the REST1 and CUFF1 scans.

First, we filter the table, keeping only the amplitudes from a single model. We'll also convert the components into a factor, for later use during modeling. At this point, the dataset is small enough that it's easy to handle in memory, and so we'll also [collect](https://dplyr.tidyverse.org/reference/compute.html) it. 

```{r}

run1_amplitudes <- amplitude |>
  filter(model == "2.1", run == 1) |>
  select(-model, -run, -ses) |>
  collect() |>
  mutate(component = factor(component))

head(run1_amplitudes)
```

To get a very high-level overview of the data, generate the [MA plot](https://en.wikipedia.org/wiki/MA_plot). Let's exclude the "red" scans to determine whether those are driving the difference.

```{r}
red_run1_scans <- read_tsv(dir_ls("data/scans", glob = "*tsv"), na = "n/a") |>
  filter(rating == "red", stringr::str_detect(filename, "cuff_run-0[12]")) |>
  mutate(
    sub = stringr::str_extract(filename, "[[:digit:]]{5}") |> as.integer()
  ) |>
  select(sub)

```

::: {#fig-falff}

```{r}
#| fig-height: 15

do_ma_plot <- function(.d) {
  .d |>
    pivot_wider(names_from = task, values_from = fALFF) |>
    na.omit() |> # not every participant has a REST1 and CUFF1
    mutate(a = (log2(rest) + log2(cuff)) / 2, m = log2(cuff) - log2(rest)) |>
    ggplot(aes(x = a, y = m)) +
    facet_wrap(~component, scales = "free", nrow = 15) +
    geom_point(alpha = 0.2) +
    geom_hline(yintercept = 0, color = "blue") +
    geom_smooth(method = "lm", color = "red") +
    xlab("Average Log Intensity") +
    ylab("CUFF1 - REST1 Log Intensities") +
    theme(strip.background = element_blank(), strip.text = element_blank())
}

run1_amplitudes_without_red <- run1_amplitudes |>
  anti_join(red_run1_scans, by = join_by(sub))

run1_amplitudes |> do_ma_plot()

```

Fractional Amplitude of Low Frequency Fluctuations. Each panel is a distinct component. The x-axis is a mean average, and the y-axis is a ratio of the two fALFFs. The 

:::

For the most part, there is not a substantially large difference in the log intensities. However, it may be interesting that there appears to be a trend such that amplitudes are somewhat lower in the CUFF as compared to REST conditions. 


```{r}
lme4::lmer(
  log(fALFF) ~ task + (1 | sub) + (1 | component),
  run1_amplitudes_without_red
) |>
  summary()
```

So, indeed, there is some evidence for higher amplitudes during REST1 as compared with CUFF1. Please note that a more careful analysis should include other factors like scanning site, participant age, or motion. 

## Considerations While Working on the Project

### Variability Across Scanners

{{< include _snippets/mri-scanner-variability.qmd >}}

### Data Quality

{{< include _snippets/mri-qc.qmd >}}

### Citations

If you use these products in your analyses, please cite the relevant papers written by members [TReNDS](https://trendscenter.org/data/).
